---
title: "R Notebook"
output: html_notebook
---

## This is code to read in a table of facilities in India and extract annual PM data from a raster.  

## This code uses two packagfes to perform the spatial operations. Here is a [good resource on the topic](https://geocompr.robinlovelace.net/spatial-data-operations.html).


First, load the two libraries
```{r}
library(sf); library(raster)
```

Next read in the csv of the facilities and take a look at it
```{r}
# read in the facilities csv
fac <- read.csv("F:/PM2.5_Data/Facilities.csv")

# view the facilities table
fac


```

Right now, this is just a table (even thought it contains Lat and lon) and we need to make it a spatial object in order to perform spatial operations on it. NOTE: `crs` is the coordinate reference system. [This one](http://spatialreference.org/ref/epsg/wgs-84/) is one commonly used by GPS.

```{r}
# turn the table into a spatial object
fac_sp <- st_as_sf(fac, coords = c("Longitude", "Latitude"), crs = 4326) 

# view that object
plot(fac_sp, axes = T)
```
Now we have our points and we need to bring in the raster data. 

```{r}
# read in the raster 
dimaq2016 <- raster("F:/PM2.5_Data/GBD2017_PM_2016.tif")
plot(dimaq2016)

```

Lastly, we extract the values in the raster where we have facilities and give those PM values to the corresponding facility. 

```{r}
extract(x = dimaq2016, y = fac_sp)
```

That function returns a vector of the PM values at each facility, and we'll just use that to populate a new column in the facility spatial object. And let's view the 2016 values at our facilities.

```{r}
fac_sp$dimaq2016 <- extract(x = dimaq2016, y = fac_sp)

plot(fac_sp["dimaq2016"], axes=T)

```

Now let's write a function to make the computer do this for all the years for which we have rasters. 

```{r}

# define the function which reads in the raster for a given year and does the extraction
extractDIMAQ <- function(year){
  
  rasterName <- paste0("F:/PM2.5_Data/GBD2017_PM_",year,".tif")
  dimaq <- raster(rasterName)
  extract(x = dimaq, y = fac_sp)

}

# define years
years <- list("2015","2016","2017")

# apply our new functino to each year in your list of years
extractions <- lapply(years, extractDIMAQ)
# it returns the result of each year's extraction as an entry in a list of the extractions

# create new variables for each year in the original 
fac_sp$year2015 <- unlist(extractions[1])
fac_sp$year2016 <- unlist(extractions[2])
fac_sp$year2017 <- unlist(extractions[3])

fac_sp

```

Let's look at the PM values at facilities over the years.
```{r}

plot(fac_sp[c("year2015","year2016","year2017")], axes=T)

```

Finally, let's get a csv out of this taht we can use in other analyses. We have to first remove the spatial information from our facilities spatial object to make it into a table.  

```{r}

fac_sp$geometry <- NULL

write.csv(fac_sp, "F:/PM2.5_Data/Facilities_w_PM.csv")

```

